---
title: "data cleaning"
output:
  pdf_document: default
  html_document: default
date: "2023-10-21"
---
```{r}
housing = read.csv('data1.csv')
housing
```

```{r}
dim(housing)
```
```{r}
# Assuming 'housing' is a data frame in R
summary_data <- summary(housing)
summary_data <- t(summary_data)
print(summary_data)

```
```{r}
# Assuming 'housing' is a data frame in R
str(housing)


```

Converting Datatype Variables
I will need to convert the datatypes for the above datatypes to their appropriate classification. Some columns such as SALE PRICE is considered an object when it should be an float. And other columns such as NEIGHBORHOOD should be a categorical datatype rather than an object.

This dataset only contains object and integer dtypes.

```{r}
# Assuming 'housing' is a data frame in R
object_columns <- sapply(housing, class) == "character"
print(names(housing[object_columns]))

```
```{r}
# Assuming 'housing' is a data frame in R
integer_columns <- sapply(housing, class) == "integer"
print(names(housing[integer_columns]))

```

```{r}
# Convert objects to categorical variables
# Check if the columns exist
for (colname in obj_cats) {
  if (colname %in% colnames(housing)) {
    # Column exists, proceed with conversion
  } else {
    cat(paste("Column", colname, "does not exist in the 'housing' data frame.\n"))
  }
}


```

```{r}
# Convert numeric to categorical variables
# Check if the columns exist
for (colname in num_cats) {
  if (colname %in% colnames(housing)) {
    # Column exists, proceed with conversion
  } else {
    cat(paste("Column", colname, "does not exist in the 'housing' data frame.\n"))
  }
}

```


```{r}
# Assuming 'housing' is a data frame in R
str(housing)  # Provides information about data types and a summary of the data

# To get a summary of the data frame
summary(housing)  # Provides summary statistics for numeric columns

# To count non-missing values in each column
colSums(!is.na(housing))

```

```{r}
head(housing)

```
```{r}
str(housing)
```

```{r}
# Assuming 'housing' is a data frame in R
housing$BOROUGH <- as.character(housing$BOROUGH)  # Ensure 'BOROUGH' is character type

# Use ifelse to replace numerical indicators with borough names
housing$BOROUGH <- ifelse(housing$BOROUGH == "1", "Manhattan",
                          ifelse(housing$BOROUGH == "2", "Bronx",
                          ifelse(housing$BOROUGH == "3", "Brooklyn",
                          ifelse(housing$BOROUGH == "4", "Queens",
                          ifelse(housing$BOROUGH == "5", "Staten Island", housing$BOROUGH)))))

```

missing value 

```{r}
# Assuming 'housing' is a data frame in R
# Use the apply function with is.na to check for missing values in each column
missing_counts <- apply(housing, 2, function(x) sum(is.na(x)))

# The missing_counts variable now contains the count of missing values for each column
print(missing_counts)

```


```{r}
# Assuming 'housing' is a data frame in R
# Use the apply function with is.na to check for missing values in each column
missing_counts <- apply(housing, 2, function(x) sum(is.na(x)))

# Calculate the percentage of missing values for each column
percentage_missing <- (missing_counts / nrow(housing)) * 100

# The percentage_missing variable now contains the percentage of missing values for each column
print(percentage_missing)

```


```{r}
housing
```

theres no data in EASE.MENT AND APARTMENT.NUMBER colomns so we drop this

```{r}
# Assuming 'housing' is a data frame in R
housing$EASEMENT <- NULL
housing$APARTMENT.NUMBER = NULL
```
```{r}
head(housing)
```

```{r}
# Assuming 'housing' is a data frame in R
# Load the graphics package
library(graphics)

# Create a matrix of missing values (1 for missing, 0 for not missing)
missing_matrix <- ifelse(is.na(housing), 1, 0)

# Set up the plot
heatmap(missing_matrix, 
        Colv = NA, 
        Rowv = NA, 
        col = c("red", 'green'),  # Specify colors for missing (grey) and non-missing (white)
        xlab = "Columns", 
        ylab = "Rows", 
        main = "Missing Values Heatmap")

```

```{r}
housing$EASE.MENT = NULL

```


drop null value cause theres huge data in this dset

```{r}


```



```{r}
str(housing)
```
```{r}
housing

```

```{r}
# Assuming 'housing' is a data frame in R
# Load the graphics package
library(graphics)

# Create a matrix of missing values (1 for missing, 0 for not missing)
missing_matrix <- ifelse(is.na(housing), 1, 0)

# Set up the plot
heatmap(missing_matrix, 
        Colv = NA, 
        Rowv = NA, 
        col = c("red", 'green'),  # Specify colors for missing (grey) and non-missing (white)
        xlab = "Columns", 
        ylab = "Rows", 
        main = "Missing Values Heatmap")
```



```{r}
# Assuming 'housing' is a data frame in R
# Remove rows with missing values
data <- na.omit(housing)

# Save the cleaned data to a CSV file
write.csv(data, file = "data.csv", row.names = FALSE)

```


```{r}
# Assuming 'housing' is a data frame in R
# Load the graphics package
library(graphics)

# Create a matrix of missing values (1 for missing, 0 for not missing)
missing_matrix <- ifelse(is.na(data), 0, 1)

# Set up the plot
heatmap(missing_matrix, 
        Colv = NA, 
        Rowv = NA, 
        col = c("red", 'green'),  # Specify colors for missing (grey) and non-missing (white)
        xlab = "Columns", 
        ylab = "Rows", 
        main = "Missing Values Heatmap")
```

